### 2020-06-06 doc/ms/GEBv1/WoS DB publication/WoS DB cleaning/Reference list WoS.Rmd
single project

---
title: "References WoS"
author: "Eduardo Fernández Pascual"
date: "17/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
library(tidyverse); library(kableExtra)
load("C:/EFP/Trabayu/GitHub/sylvanseeds/data/WoSDB.RData")
References %>% select(Reference) %>% arrange(Reference) %>% kable(caption = "Appendix 1 – Data sources. List of references used to build the germination database.", booktabs = TRUE) %>% kable_styling(latex_options = c("striped", "hold_position"))
```

Refs