### 2020-06-06 README.Rmd
readme revised

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![Sylvan Seeds](data/2017070313 Blue Ridge Parkway small.jpg)

# Sylvan Seeds

*Sylvan Seeds* is a dataset of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >600 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

## Contents

The following materials are available in this repository:

* [The dataset file itself is `.csv` format](https://github.com/efernandezpascual/sylvanseeds/blob/master/results/Supplementary%20material%206%20-%20Database.csv)
* [The Sylvan Seeds `shiny` app for visualizing the dataset](https://sylvanseeds.shinyapps.io/sylvanseeds/)
* [The data paper describing the collection and fields of the dataset](https://github.com/efernandezpascual/sylvanseeds/blob/master/doc/last/last.md), which is in preparation.

## Citation

Please cite the dataset as `CITATION`