### 2020-09-15 README.Rmd
JVS minor revision

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![SylvanSeeds](data/2017070313 Blue Ridge Parkway small.jpg)

# SylvanSeeds, a seed germination database for temperate deciduous forests

This repository stores all the information related to the publication *SylvanSeeds, a seed germination database for temperate deciduous forests*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.

*SylvanSeeds* is a dataset of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >500 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in this repository:

* `data` Original raw files including the primary result of the search in the Web of Science, and search notes.
* `doc` Files to create the manuscript using `Rmarkdown`.
* `results` Output of the `R` scripts, including the clean database and files needed to create the manuscript.
* `src` Scripts in `R` language used to clean the raw data, perform the analyses of the manuscript, and build the `SylvanSeeds` app.
* [The database itself in `.csv` format](https://github.com/efernandezpascual/sylvanseeds/blob/master/results/Supplementary%20material%206%20-%20Database.csv)
* [The SylvanSeeds `shiny` app for visualizing the dataset](https://sylvanseeds.shinyapps.io/sylvanseeds/)
* [The report paper describing the collection and fields of the dataset](https://github.com/efernandezpascual/sylvanseeds/blob/master/doc/manuscript.md), published in the Journal of Vegetation Science.

## Citation

Please cite the dataset as

> Fernández-Pascual, E (2020) SylvanSeeds, a seed germination database for temperate deciduous forests. Journal of Vegetation Science, in press.