### 2020-09-15 README.Rmd
3rd submission to JVS 🌳🍋

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

![SylvanSeeds](data/2017070313 Blue Ridge Parkway small.JPG)

# SylvanSeeds, a seed germination database for temperate deciduous forests

[*SylvanSeeds*](https://efernandezpascual.github.io/home/sylvanseeds.html) is a database of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >500 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

This repository stores all the information related to the publication *SylvanSeeds, a seed germination database for temperate deciduous forests*, including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript. It also stores the code of the *SylvanSeeds* `shiny` app, that can be used to visualize the database.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in this repository:

* `data` Original raw files including the primary result of the search in the Web of Science, and search notes.
* `doc` Files to create the manuscript using `Rmarkdown`.
* `results` Output of the `R` scripts, including the clean database and files needed to create the manuscript.
* `src` Scripts in `R` language used to clean the raw data, perform the analyses of the manuscript, and build the `SylvanSeeds` app.

## Citation

Please cite `SylvanSeeds` as:

> Fernández-Pascual, E (2020) SylvanSeeds, a seed germination database for temperate deciduous forests. Journal of Vegetation Science, in press.