### 2020-11-12 README.Rmd
revised folder structure 🍂

---
title: "README.md"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4081377.svg)](https://doi.org/10.5281/zenodo.4081377)

![SylvanSeeds](2017070313 Blue Ridge Parkway small.jpg)

# SylvanSeeds, a seed germination database for temperate deciduous forests

[*SylvanSeeds*](https://efernandezpascual.github.io/home/sylvanseeds.html) is a database of seed germination ecology for the temperate broadleaf and mixed forest biome of the northern hemisphere. It contains data on >300 plant species, obtained from >500 published references. The data is in the form of germination proportions recorded under experimental combinations of temperature, light, scarification and stratification.

This repository stores all the information related to the publication [*SylvanSeeds, a seed germination database for temperate deciduous forests*](https://doi.org/10.1111/jvs.12960), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the [manuscript](https://github.com/efernandezpascual/sylvanseeds/blob/master/doc/manuscript.md). It also stores the code of the *SylvanSeeds* `shiny` app, that can be used to visualize the database.

## Contents

This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.

The following materials are available in this repository:

* `data` Primary data files including the SylvanSeeds database in csv format and Boolean search string used in the Web of Science.
* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/sylvanseeds/blob/master/doc/manuscript.md) using `Rmarkdown`.
* `results` Output of the `R` scripts, including files needed to create the manuscript and run the app.
* `src` Scripts in `R` language used to clean data, perform analyses, and build the `SylvanSeeds` app data.

## Citation

Please cite `SylvanSeeds` as:

> Fernández-Pascual E (2021) [SylvanSeeds, a seed germination database for temperate deciduous forests.](https://doi.org/10.1111/jvs.12960) *Journal of Vegetation Science.*